---
title: "R Notebook"
output: html_notebook
---
Going through the tutoral of drawing heatmap with R. From Dave Tang's blog:
https://davetang.org/muse/2010/12/06/making-a-heatmap-with-r/

Using the TagSeqExample.tab file from DESeq.

# 1. Using the basic `heatmap`function:
```{r}
#install DESeq if necessary
#source("http://bioconductor.org/biocLite.R")
#biocLite("DESeq")

# Load DESeq:
library("DESeq")

# Import the TagSeqExample.tab file from DESeq:
## Finds the full file names of files in DEseq packages:
example_file <- system.file ("extra/TagSeqExample.tab", package="DESeq")
## Load the data: (He delete the rownames.)
data <- read.delim(example_file, header=T, row.names="gene")

# Check the data:
head(data)
nrow(data)
## Check distribution of row sums
quantile(rowSums(data))
   


# Create a workable set
data_subset <- data[rowSums(data)>50000,]
nrow(data_subset)
## Create the data matrix for heatmap:
data_matrix <- data.matrix(data_subset)

# Generate the heatmap
heatmap(data_matrix)
```

# 2. Using the `heatmap.2`: 
```{r}
# install if necessary
# install.packages("gplots")
library(gplots)
heatmap.2(data_matrix)

# Center the values and scaled in row direction:
heatmap.2(data_matrix,scale="row")

# Using a red and blue colour scheme without traces and scaling each row:
library("RColorBrewer")
heatmap.2(data_matrix,col=brewer.pal(11,"RdBu"),scale="row", trace="none")

```

# 3. Using pheatmap package (basic):
From the other assay in Tang's blog.
https://davetang.org/muse/2018/05/15/making-a-heatmap-in-r-with-the-pheatmap-package/

```{r}
# load package
library(pheatmap)

# The default (w/o scaling)
pheatmap(data_matrix)

# Scale in row direction:
pheatmap(data_matrix, scale="row")

```


# 4. Using pheatmap package (advanced):
The power of pheatmap: the ability to add annotations to the rows and columns!
```{r}
# Using `dendextend` to creat hierarchical clustering for deciding how to color the plot: (You can also extract the information directly from pheatmap!)
my_hclust_gene <- hclust(dist(data_matrix), method = "complete")
## Load package
library(dendextend)
## Make the dendrogram: (clustering in the same manner as performed by pheatmap)
as.dendrogram(my_hclust_gene) %>%
  plot(horiz = TRUE)

# Define 2 sub-clusters by cutting the tree: (the cluster number: k=2)
my_gene_col <- cutree(tree = as.dendrogram(my_hclust_gene), k = 2)
my_gene_col
## Annotate the genes by the sub-clusters:
my_gene_col <- data.frame(cluster = ifelse(test = my_gene_col == 1, yes = "cluster 1", no = "cluster 2"))
head(my_gene_col)


# Add more annotations:
## Add some random annotations:
set.seed(1984)
my_random <- as.factor(sample(x = 1:2, size = nrow(my_gene_col), replace = TRUE))
my_gene_col$random <- my_random
head(my_gene_col)
## Add the column annotations:
my_sample_col <- data.frame(sample = rep(c("tumour", "normal"), c(4,2)))
row.names(my_sample_col) <- colnames(data_matrix)


# Create the heatmap with all annotations!
pheatmap(data_matrix, annotation_row = my_gene_col, annotation_col = my_sample_col)


# Other useful feature: breaking up the heatmap!
pheatmap(data_matrix,
         annotation_row = my_gene_col,
         annotation_col = my_sample_col,
         cutree_rows = 2,
         cutree_cols = 2)

```

# 5. Others
Extract the hierarchical clustering from pheatmap!
```{r}
# use silent = TRUE to suppress the plot
my_heatmap <- pheatmap(data_matrix, silent = TRUE)
 
# results are stored as a list
class(my_heatmap)
names(my_heatmap)

my_heatmap$tree_row %>%
  as.dendrogram() %>%
  plot(horiz = TRUE)
```

Saving the heatmap as a PNG file:
```{r}
save_pheatmap_png <- function(x, filename, width=1200, height=1000, res = 150) {
  png(filename, width = width, height = height, res = res)
  grid::grid.newpage()
  grid::grid.draw(x$gtable)
  dev.off()
}
 
#save_pheatmap_png(my_heatmap, "my_heatmap.png")
```

Note:
> I acciednetally found pheatmap can accept data.frame.


